\vskip-0.6in Trade Agreements in the Shadow of Lobbying\protect\footnote{Formerly circulated under the title ``Trade Agreements, Lobbying and Separation of Powers''}

### 2016-02-23 draft.tex
remove bold / color commands from first round revision

\begin{quote}
{This paper presents a model of international trade agreements in which the executive branches of each government negotiate agreements while the legislative branches, subject to political pressure from firms, can disrupt them. Lobbying is in the style of Grossman and Helpman (1994) with a new feature: all actors face uncertainty arising from the complexity of the legislative process. I demonstrate that the lower the executives set tariffs in a trade agreement, the more effort lobbies put forth to prevent its ratification. Thus trade agreements act as a domestic political commitment device: executives set relatively high tariffs to discourage lobbying and increase the chance that the agreement will be ratified. \textbf{The model sheds light on the empirical puzzle surrounding governments' welfare weights in the Grossman and Helpman (1994) model and can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.}

@@

\textbf{Much of trade policy is determined in the context of international agreements that must be ratified by some domestic body. This paper seeks to shed light on such a policy-making process by incorporating Protection-for-Sale (Grossman and Helpman 1994, hereafter PFS) style lobbying into a model of trade agreements in which the executive branches of the governments set tariff levels in anticipation of political pressure upon a ratifying body, assumed here to be a legislature. This political pressure may be interpreted either as lobbying against ratification of the trade agreement or lobbying to break the agreement once in force. Important realism is achieved not only through the addition of the ratification constraint but also by the incorporation of political uncertainty into the model.}

\textbf{Assuming that lobbies represent only import-competing interests, lobbies in this environment would like to prevent or derail a trade agreement that would reduce their tariff protection. They work toward this end by exerting effort to convince their legislatures to vote against the trade agreement but face some uncertainty about whether their efforts will result in a voting outcome consistent with their interests. I show that the probability that the trade agreement is derailed increases in lobbying effort and decreases in the trade agreement tariff. Further, the lobby decreases its effort as the tariff it would face under the trade agreement rises: it has less to lose the higher is the trade agreement tariff. When forming the trade agreement tariff, the executives therefore trade off the welfare they receive under the agreement with the probability that the agreement actually remains in force. Ultimately, I find that observed lobbying effort may often be low because the executive branches of the governments set trade agreement tariffs sufficiently high to reduce political pressure on the legislatures. There are other ways in which realism is gained when political uncertainty is added to the model: for instance, the model rationalizes the four-year delay and eventual ratification of the U.S.-Korea Free Trade Agreement.}

\textbf{To see how this works,} consider first the case of political certainty, and assume that the executive is more pro-social than the legislature, as has been the case throughout the post-war period in the United States.\footnote{The standard assumption that the executive is less protectionist than the legislature is a special case of the finding that susceptibility to special interests generally declines with constituency size. One simple illustration from the realm of trade policy is the following: a legislator whose district has a large concentration of a particular industry does not take into account the impact of tariffs on the welfare of consumers in other districts, while the executive, whose constituency encompasses the whole country, will internalize these diffuse consumption effects. See Lohmann and O'Halloran (1994) for a detailed argument and Grossman and Helpman (2005) for a formal model that demonstrates the point for the legislature.\label{fn:ga_l_e}} For any trade-agreement tariff set by the executives, the lobby can calculate the effort level required to prevent the agreement from being ratified: it must shift the preferences of the median legislator just far enough so that he will choose the non-cooperative tariffs over the trade agreement tariffs. Because the lobby will pay this price so long as its benefit outweighs the cost, a trade war will occur unless the executive sets the trade-agreement tariff so that lobbying is not worthwhile.\footnote{As will be made clear below, non-ratification is equivalent to breaking a trade agreement once in place; likewise, the non-cooperative tariffs that result from failing to ratify the trade agreement are the same as the tariffs that would arise in trade war, so this model applies to both situations.}

Thus, any trade agreement under political certainty involves a tariff high enough to disengage the lobby completely; that is, even welfare-maximizing executives set high tariffs and induce zero contributions in equilibrium because they set policy in the shadow of a ratification decision influenced by lobbying. Careful modeling of the political process demonstrates that \textbf{the relationship between deadweight loss calculations and the tariffs that are provided in a trade agreement may be mediated in important ways by the ratification constraint.}

The fact that \textit{zero} contributions are predicted in equilibrium is not particularly satisfying. The addition of political uncertainty provides the missing realism, smooths the results and delivers additional intuition. I assume all actors are uncertain about the weight the median legislator in a non-unitary legislature places on the profits of the lobbying industry. \textbf{This is meant to represent the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote. For instance, perhaps} a small number of lobbied legislators cannot deliver the votes of their non-lobbied colleagues with certainty. \textbf{One can think of this uncertainty as an additional influence on the legislature's preferences about which all parties---executives, legislators and lobbies alike---are uncertain.}

In this case \textbf{of political uncertainty}, I show that the legislatures disrupt trade agreements with a higher probability and set higher trade war tariffs when lobbying effort increases. Because the lobbies respond to higher trade agreement tariffs by decreasing effort (and therefore the probability the agreement will fail to be ratified or will be broken), the executives must trade off the level of social welfare derived while a trade agreement is in force---since they prefer low tariffs---and the chances that the agreement will remain in force.

@@

\textbf{The addition of a rich structure of government}, tailored carefully to the trade policy making process, also helps shed light on the empirical puzzle surrounding the PFS model. Empirical tests of the model\footnote{cfr. Goldberg and Maggi (1999); Gawande and Bandyopadhyay (2000); Mitra, Thomakos and Ulubasoglu (2002); McCalman (2004).\label{fn:gh_lobby}} have consistently found the weight governments place on social welfare to be many times that which they place on lobbying effort, while estimates indicate that the deadweight losses caused by trade distortions are several orders of magnitude larger than lobbying expenditures.\footnote{Feenstra (1992) assembles estimates from the mid-1980s that place a floor of $\$$8 billion per year on the deadweight losses from protection, whereas Bombardini and Trebbi (2012) find that total annual lobbying expenditure on trade issues in 1999-2001 when this data first became available was about $\$$200 million.} This raises the question: how could governments value social welfare so highly, yet grant these quantities of protection at such a low price?

@@

In the non-unitary government model, these estimates and the earlier-cited stylized facts become consistent with each other. The ``case of the missing contributions,'' as Gawande and Hoekman (2006) refer to it, turns out to be well explained by an equilibrium phenomenon arising from the agenda-setting power of the executive branch. In particular, \textbf{in this model, there is no inconsistency for the approximately 15$\%$ of sectors that receive protection but apparently put forth no lobbying effort. In this framework, even a welfare-maximizing executive and lobbying effort at zero can be reconciled with high tariff levels.}

@@

Another related literature considers the impact of exogenously-determined political uncertainty on the potential for trade cooperation. These studies (e.g. Feenstra and Lewis (1991); Milner and Rosendorff (2001); Bagwell and Staiger (2005); Beshkar (2010); \textbf{Horn, Maggi and Staiger (2010); Beshkar and Bond (2012); Amador and Bagwell (2013)}) derive various implications for the design of international trade agreements using a Baldwin (1987)-style government welfare function with exogenous shocks to the political-pressure parameter. As explained in the next section, the model presented here is intended to be a bridge between this literature and that discussed above that focuses on endogenously-determined political economy weights. To preview, this model is closest to that of Bagwell and Staiger (2005), with two main changes: in place of their unitary government signing a trade agreement and having different preferences ex-ante and ex-post, this model has two branches of government with differing preferences, where---in the case of the legislature---the political economy weights are determined endogenously.

\textbf{A very recent literature has begun to study the effects of trade agreements for reducing the uncertainty that firms (Handley 2014 and Handley and Lim\~{a}o 2015) and governments (Lim\~{a}o and Maggi 2015) face.}

@@

\subsection{The Basic Setup}
\label{sec:basic}
\textbf{I begin by describing the basic economic setting within which trade occurs. It is a three-good model with two countries: home (no asterisk) and foreign (asterisk). In each country, preferences are linear in good $N$, which is denoted the numeraire, while the demand functions for $X$ and $Y$ are assumed strictly decreasing and twice continuously differentiable. Further, the demand functions for $X$ and $Y$ are taken to be identical and denoted $D(P_i)$ in home and $D(P_i^*)$ in foreign. $P_i$ denotes the home price of good $i \in \left\{N,X,Y\right\}$ and $P_i^*$ denotes the foreign price of good $i$.}

\textbf{Good $N$ is produced with labor alone with constant returns to scale technology so that $Q_N = l_N$. I assume the aggregate labor supply $l$ is large enough to ensure that the output of good $N$ is large enough to guarantee balanced trade.} The supply functions for good $X$ are $Q_X(P_X)$ and $Q_X^*(P_X^*)$ and are assumed strictly increasing and twice continuously differentiable for all prices that elicit positive supply. I also assume $Q_X^*(P_X) > Q_X(P_X)$ for any such $P_X$ so that the home country is a net importer of good $X$. The production structure for good $Y$ is taken to be symmetric, with both demand and supply such that the economy is separable in goods $X$ and $Y$. It is assumed that the production of goods $X$ and $Y$ requires, \textbf{in addition to labor,} the possession of a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factors is tied to the price of the good in whose production their factor is used.

For simplicity, I assume each government's only trade policy instrument is a specific tariff on its import-competing good: the home country levies a tariff $\tau$ on good $X$ while the foreign country applies a tariff $\tau^*$ to good $Y$. Local prices are then $P_X = P_X^W + \tau$, $P_X^* = P_X^W$, $P_Y = P_Y^W$ and $P_Y^* = P_Y^W + \tau^*$ where a $W$ superscript indicates world prices. Equilibrium prices are determined by the market clearing conditions
$$M_X(P_X)= D(P_X)-Q_X(P_X) = Q_X^*(P_X^*) - D(P_X^*) = E_X^*(P_X^*)$$
$$E_Y(P_Y)=Q_Y(P_Y)-D(P_Y) = D(P_Y^*)-Q_Y^*(P_Y) = M_Y^*(P_Y^*)$$
where $M_X$ are home-county imports and $E_X^*$ are foreign exports of good $X$ and $E_Y$ are home-county exports and $M_Y^*$ are foreign imports of good $Y$. \textbf{Given the production technology and no trade taxes, the price of the numeraire is equal to one in both countries and on the world market.}

$P_X^W$ and $P_Y^W$ are decreasing in $\tau$ and $\tau^*$ respectively, while $P_X$ and $P_Y^*$ are increasing in the respective domestic tariff. This gives rise to a standard terms-of-trade externality. As profits and producer surplus (identical in this model) in a sector are increasing in the price of its good, profits in the import-competing sector are also increasing in the domestic tariff. This economic fact, combined with the assumptions on specific factor ownership, is what motivates political activity \textbf{by import-competing lobbies}.

\textbf{I now turn to describing the strategic players in the game.} In order to focus attention on protectionist political forces, I assume that only the import-competing industry in each country is politically-organized and able to lobby and that it is represented by a single lobbying organization (see Footnote~\ref{fn:lobby} for discussion of additional lobbies). Each country's government is composed of two branches: an executive who can conclude trade agreements and a legislature that has final say on trade policy. \textbf{To summarize}, the political process is modeled as involving three players in each country: the lobby, the executive, and the legislature.

\textbf{Figure~\ref{fig:ext} illustrates the timing of the game from the perspective of the home country.}\footnote{\textbf{This extensive form is for the case where the foreign legislature does not break the trade agreement.}} First, the executives set trade policy cooperatively within the context of an international agreement, \textbf{that is they choose the trade agreement tariffs $\bta = \left(\tau^a,\tau^{*a}\right)$}. After the trade agreement is concluded, the lobbies attempt to persuade the legislators in their respective countries to vote ``no'' on the ratification of the agreement; or alternatively, if there is no ratification hurdle, to break the trade agreement.\footnote{As will be seen shortly, these two interpretations are equivalent in this setting because tariffs in a ``trade war'' are assumed to be set non-cooperatively in the same manner as if there were no trade agreement. For simplicity, I refer to this as the ``break'' decision throughout.} \textbf{They do this by choice of lobbying effort $e_b$, which can take on any non-negative value.}

\textbf{Next, nature determines with probabilities $o$ and $o^*$ whether each legislature will have the opportunity to take a vote to break the agreement. I assume that these probabilities are less than one-half and mutually exclusive.}\footnote{\textbf{While this assumption is made for purposes of exposition and tractability, one can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.}} \textbf{This allows me to focus on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.}\footnote{\textbf{When the legislatures in both countries have an opportunity to break the agreement simultaneously, an incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.}}

\textbf{After this}, the uncertainty about the the median legislator's identity is resolved. This is modeled as the realization of a random variable $\ta_b$, the value of which all players observe simultaneously. The ``break stage'' concludes with the legislature making a choice to abide by the agreement or to break the agreement.

@@

\begin{figure}
	\begin{center}
		\input{game-tree.tex}
	\end{center}
	\caption{\textbf{Extensive Form\label{fig:ext}}}
\end{figure}

\subsection{Preferences}
\label{sec:pref}
As the economy is fully separable and the economic and political structures are symmetric, I focus throughout the paper on the home country and the $X$-sector. The details are analogous for $Y$ and foreign.\footnote{\textbf{These are the relevant sectors for analysis because they are the sectors in which there are lobbies.}}

I specify the welfare function of the home legislature as
\begin{equation}
  W_{\mathit{ML}} = \mathit{CS}_X(\tau) + \ga(e,\ve) \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
  \label{eq:ml}
\end{equation}
where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{\textbf{Labor income $l$ could also be included in both median legislator and executive welfare. I omit it because its inclusion alters none of the results and only serves to complicate the exposition.}} I model the decisions of the legislature as being taken by a median legislator \textbf{(more on this in Section~\ref{sec:choices}). Notice that} the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.

Given its expectations and the legislature's preferences, the home lobby chooses its lobbying effort ($e_b$ to influence the break/ratification decision, which for simplicity, I will refer to henceforth simply as ``break,'' and $e_n$ to influence the trade war tariff) to maximize the welfare function:
\begin{equation}
  {\color{blue} \expect \left[U_L \right]} = \Pr\left[ \text{TradeWar}(\bta,e_b,e_n) \right] \left[ \pi(\tau^{\mathit{n}}) - e_n \right] + \Pr\left[ \text{TradeAgreement}(\bta,e_b,e_n) \right] \pi(\tau^a) - e_b
  \label{eq:lv}
\end{equation}
where $\pi(\cdot)$ is the current-period profit and $\tau^a \ (\tau^\mathit{n})$ is the home country's tariff on the import good under a trade agreement (war). I use the convention throughout of representing a vector of tariffs for both countries $(\tau,\tau^*)$ as a single bold $\bm{\tau}$.

In the first stage \textbf{of the game}, the executives choose the trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ via a negotiating process that I assume to be efficient. In this symmetric environment, it therefore maximizes the expected joint payoffs:
\begin{multline}
  {\color{blue}\expect \left[W_E(\bta) \right] + \expect \left[ W_E^{*}(\bta) \right]} = \Pr\left[ \text{TradeWar}(\bta) \right] \left[W_E(\bm{\tau}^{\mathit{n}}) + W_E^*(\bm{\tau}^{\mathit{n}}) \right] + \\ \Pr\left[ \text{TradeAgreement}(\bta) \right] \left[W_E(\bta) + W_E^*(\bta) \right]
  \label{eq:jv}
\end{multline}

@@

The home executive's welfare is specified as follows:
\[
  W_E = \mathit{CS}_X(\tau) + \ga_E \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
\]
Note that this is identical to the welfare function for the legislature aside from the weight $\ga_E$ the executive places on the profits of the import industry, which is not a function of lobbying effort. \textbf{Section~\ref{sec:choices} contains an in-depth discussion of these modeling choices.}

\subsection{Information and Equilibrium Selection}
\label{sec:info}
\textbf{I examine a simple class of equilibria that have the following features.}

\textbf{First, on information.} Although uncertainty is present, information about it is symmetric. However, in line with the literature, I assume the lobby's contribution is not observable to the foreign legislature. The implication is that the lobby can directly influence only the home legislature, and so the influence of one country's lobby on the other country's legislature occurs only through the tariffs selected.\footnote{cfr. Grossman and Helpman (1995b), page 685.} \textbf{Given the game's structure and the assumption that lobbying effort is not observable across international borders, players in the same country can take advantage of more information than those who are in different countries. That is, the solution concept employed here is perfect public equilibrium (PPE).}

\textbf{Second, whenever there is a possibility of multiple equilibria, I will focus on the one that maximizes the welfare of the executives. }

\textbf{Finally, I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.}\footnote{Buzard (\textbf{2015}) considers a similar model in a repeated-game context and can therefore remove this restriction.}

@@

In modeling the objective of the legislature \textbf{(see Equation~\ref{eq:ml}), I have introduced} a variant of the general political support function introduced by Hillman (1982). While Hillman is credited with first proposing that policy makers choose tariffs in order to maximize their own political support, the PFS model contributed the explicit specification that lobbying contributions enter the political support function with constant marginal utility as $C + aW$.

@@

\noindent \textbf{All proofs are in the Appendix.}

>>>>

\begin{quote}
{This paper presents a model of international trade agreements in which the executive branches of each government negotiate agreements while the legislative branches, subject to political pressure from firms, can disrupt them. Lobbying is in the style of Grossman and Helpman (1994) with a new feature: all actors face uncertainty arising from the complexity of the legislative process. I demonstrate that the lower the executives set tariffs in a trade agreement, the more effort lobbies put forth to prevent its ratification. Thus trade agreements act as a domestic political commitment device: executives set relatively high tariffs to discourage lobbying and increase the chance that the agreement will be ratified. The model sheds light on the empirical puzzle surrounding governments' welfare weights in the Grossman and Helpman (1994) model and can speak to important trade policy-making events such as the four-year delay in the ratification of the U.S.-Korea Free Trade Agreement.

@@

Much of trade policy is determined in the context of international agreements that must be ratified by some domestic body. This paper seeks to shed light on such a policy-making process by incorporating Protection-for-Sale (Grossman and Helpman 1994, hereafter PFS) style lobbying into a model of trade agreements in which the executive branches of the governments set tariff levels in anticipation of political pressure upon a ratifying body, assumed here to be a legislature. This political pressure may be interpreted either as lobbying against ratification of the trade agreement or lobbying to break the agreement once in force. Important realism is achieved not only through the addition of the ratification constraint but also by the incorporation of political uncertainty into the model.

Assuming that lobbies represent only import-competing interests, lobbies in this environment would like to prevent or derail a trade agreement that would reduce their tariff protection. They work toward this end by exerting effort to convince their legislatures to vote against the trade agreement but face some uncertainty about whether their efforts will result in a voting outcome consistent with their interests. I show that the probability that the trade agreement is derailed increases in lobbying effort and decreases in the trade agreement tariff. Further, the lobby decreases its effort as the tariff it would face under the trade agreement rises: it has less to lose the higher is the trade agreement tariff. When forming the trade agreement tariff, the executives therefore trade off the welfare they receive under the agreement with the probability that the agreement actually remains in force. Ultimately, I find that observed lobbying effort may often be low because the executive branches of the governments set trade agreement tariffs sufficiently high to reduce political pressure on the legislatures. There are other ways in which realism is gained when political uncertainty is added to the model: for instance, the model rationalizes the four-year delay and eventual ratification of the U.S.-Korea Free Trade Agreement.

To see how this works, consider first the case of political certainty, and assume that the executive is more pro-social than the legislature, as has been the case throughout the post-war period in the United States.\footnote{The standard assumption that the executive is less protectionist than the legislature is a special case of the finding that susceptibility to special interests generally declines with constituency size. One simple illustration from the realm of trade policy is the following: a legislator whose district has a large concentration of a particular industry does not take into account the impact of tariffs on the welfare of consumers in other districts, while the executive, whose constituency encompasses the whole country, will internalize these diffuse consumption effects. See Lohmann and O'Halloran (1994) for a detailed argument and Grossman and Helpman (2005) for a formal model that demonstrates the point for the legislature.\label{fn:ga_l_e}} For any trade-agreement tariff set by the executives, the lobby can calculate the effort level required to prevent the agreement from being ratified: it must shift the preferences of the median legislator just far enough so that he will choose the non-cooperative tariffs over the trade agreement tariffs. Because the lobby will pay this price so long as its benefit outweighs the cost, a trade war will occur unless the executive sets the trade-agreement tariff so that lobbying is not worthwhile.\footnote{As will be made clear below, non-ratification is equivalent to breaking a trade agreement once in place; likewise, the non-cooperative tariffs that result from failing to ratify the trade agreement are the same as the tariffs that would arise in trade war, so this model applies to both situations.}

Thus, any trade agreement under political certainty involves a tariff high enough to disengage the lobby completely; that is, even welfare-maximizing executives set high tariffs and induce zero contributions in equilibrium because they set policy in the shadow of a ratification decision influenced by lobbying. Careful modeling of the political process demonstrates that the relationship between deadweight loss calculations and the tariffs that are provided in a trade agreement may be mediated in important ways by the ratification constraint.

The fact that zero contributions are predicted in equilibrium is not particularly satisfying. The addition of political uncertainty provides the missing realism, smooths the results and delivers additional intuition. I assume all actors are uncertain about the weight the median legislator in a non-unitary legislature places on the profits of the lobbying industry. This is meant to represent the idea that the lobbying and voting ``game'' that goes on within legislatures is often complex enough that none of the actors know precisely what the outcome will be---that is, no one can predict exactly which legislator will be the median and therefore what weight will be placed on import-industry profits at the time of the vote. For instance, perhaps a small number of lobbied legislators cannot deliver the votes of their non-lobbied colleagues with certainty. One can think of this uncertainty as an additional influence on the legislature's preferences about which all parties---executives, legislators and lobbies alike---are uncertain.

In this case of political uncertainty, I show that the legislatures disrupt trade agreements with a higher probability and set higher trade war tariffs when lobbying effort increases. Because the lobbies respond to higher trade agreement tariffs by decreasing effort (and therefore the probability the agreement will fail to be ratified or will be broken), the executives must trade off the level of social welfare derived while a trade agreement is in force---since they prefer low tariffs---and the chances that the agreement will remain in force.

@@

The addition of a rich structure of government, tailored carefully to the trade policy making process, also helps shed light on the empirical puzzle surrounding the PFS model. Empirical tests of the model\footnote{cfr. Goldberg and Maggi (1999); Gawande and Bandyopadhyay (2000); Mitra, Thomakos and Ulubasoglu (2002); McCalman (2004).\label{fn:gh_lobby}} have consistently found the weight governments place on social welfare to be many times that which they place on lobbying effort, while estimates indicate that the deadweight losses caused by trade distortions are several orders of magnitude larger than lobbying expenditures.\footnote{Feenstra (1992) assembles estimates from the mid-1980s that place a floor of $\$$8 billion per year on the deadweight losses from protection, whereas Bombardini and Trebbi (2012) find that total annual lobbying expenditure on trade issues in 1999-2001 when this data first became available was about $\$$200 million.} This raises the question: how could governments value social welfare so highly, yet grant these quantities of protection at such a low price?

@@

In the non-unitary government model, these estimates and the earlier-cited stylized facts become consistent with each other. The ``case of the missing contributions,'' as Gawande and Hoekman (2006) refer to it, turns out to be well explained by an equilibrium phenomenon arising from the agenda-setting power of the executive branch. In particular, in this model, there is no inconsistency for the approximately 15$\%$ of sectors that receive protection but apparently put forth no lobbying effort. In this framework, even a welfare-maximizing executive and lobbying effort at zero can be reconciled with high tariff levels.

@@

Another related literature considers the impact of exogenously-determined political uncertainty on the potential for trade cooperation. These studies (e.g. Feenstra and Lewis (1991); Milner and Rosendorff (2001); Bagwell and Staiger (2005); Beshkar (2010); Horn, Maggi and Staiger (2010); Beshkar and Bond (2012); Amador and Bagwell (2013)) derive various implications for the design of international trade agreements using a Baldwin (1987)-style government welfare function with exogenous shocks to the political-pressure parameter. As explained in the next section, the model presented here is intended to be a bridge between this literature and that discussed above that focuses on endogenously-determined political economy weights. To preview, this model is closest to that of Bagwell and Staiger (2005), with two main changes: in place of their unitary government signing a trade agreement and having different preferences ex-ante and ex-post, this model has two branches of government with differing preferences, where---in the case of the legislature---the political economy weights are determined endogenously.

A very recent literature has begun to study the effects of trade agreements for reducing the uncertainty that firms (Handley 2014 and Handley and Lim\~{a}o 2015) and governments (Lim\~{a}o and Maggi 2015) face.

@@

\subsection{The Basic Setup}
\label{sec:basic}
I begin by describing the basic economic setting within which trade occurs. It is a three-good model with two countries: home (no asterisk) and foreign (asterisk). In each country, preferences are linear in good $N$, which is denoted the numeraire, while the demand functions for $X$ and $Y$ are assumed strictly decreasing and twice continuously differentiable. Further, the demand functions for $X$ and $Y$ are taken to be identical and denoted $D(P_i)$ in home and $D(P_i^*)$ in foreign. $P_i$ denotes the home price of good $i \in \left\{N,X,Y\right\}$ and $P_i^*$ denotes the foreign price of good $i$.

Good $N$ is produced with labor alone with constant returns to scale technology so that $Q_N = l_N$. I assume the aggregate labor supply $l$ is large enough to ensure that the output of good $N$ is large enough to guarantee balanced trade. The supply functions for good $X$ are $Q_X(P_X)$ and $Q_X^*(P_X^*)$ and are assumed strictly increasing and twice continuously differentiable for all prices that elicit positive supply. I also assume $Q_X^*(P_X) > Q_X(P_X)$ for any such $P_X$ so that the home country is a net importer of good $X$. The production structure for good $Y$ is taken to be symmetric, with both demand and supply such that the economy is separable in goods $X$ and $Y$. It is assumed that the production of goods $X$ and $Y$ requires, in addition to labor, the possession of a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factors is tied to the price of the good in whose production their factor is used.

For simplicity, I assume each government's only trade policy instrument is a specific tariff on its import-competing good: the home country levies a tariff $\tau$ on good $X$ while the foreign country applies a tariff $\tau^*$ to good $Y$. Local prices are then $P_X = P_X^W + \tau$, $P_X^* = P_X^W$, $P_Y = P_Y^W$ and $P_Y^* = P_Y^W + \tau^*$ where a $W$ superscript indicates world prices. Equilibrium prices are determined by the market clearing conditions
$$M_X(P_X)= D(P_X)-Q_X(P_X) = Q_X^*(P_X^*) - D(P_X^*) = E_X^*(P_X^*)$$
$$E_Y(P_Y)=Q_Y(P_Y)-D(P_Y) = D(P_Y^*)-Q_Y^*(P_Y) = M_Y^*(P_Y^*)$$
where $M_X$ are home-county imports and $E_X^*$ are foreign exports of good $X$ and $E_Y$ are home-county exports and $M_Y^*$ are foreign imports of good $Y$. Given the production technology and no trade taxes, the price of the numeraire is equal to one in both countries and on the world market.

$P_X^W$ and $P_Y^W$ are decreasing in $\tau$ and $\tau^*$ respectively, while $P_X$ and $P_Y^*$ are increasing in the respective domestic tariff. This gives rise to a standard terms-of-trade externality. As profits and producer surplus (identical in this model) in a sector are increasing in the price of its good, profits in the import-competing sector are also increasing in the domestic tariff. This economic fact, combined with the assumptions on specific factor ownership, is what motivates political activity by import-competing lobbies.

I now turn to describing the strategic players in the game. In order to focus attention on protectionist political forces, I assume that only the import-competing industry in each country is politically-organized and able to lobby and that it is represented by a single lobbying organization (see Footnote~\ref{fn:lobby} for discussion of additional lobbies). Each country's government is composed of two branches: an executive who can conclude trade agreements and a legislature that has final say on trade policy. To summarize, the political process is modeled as involving three players in each country: the lobby, the executive, and the legislature.

Figure~\ref{fig:ext} illustrates the timing of the game from the perspective of the home country.\footnote{This extensive form is for the case where the foreign legislature does not break the trade agreement.} First, the executives set trade policy cooperatively within the context of an international agreement, that is they choose the trade agreement tariffs $\bta = \left(\tau^a,\tau^{*a}\right)$. After the trade agreement is concluded, the lobbies attempt to persuade the legislators in their respective countries to vote ``no'' on the ratification of the agreement; or alternatively, if there is no ratification hurdle, to break the trade agreement.\footnote{As will be seen shortly, these two interpretations are equivalent in this setting because tariffs in a ``trade war'' are assumed to be set non-cooperatively in the same manner as if there were no trade agreement. For simplicity, I refer to this as the ``break'' decision throughout.} They do this by choice of lobbying effort $e_b$, which can take on any non-negative value.

Next, nature determines with probabilities $o$ and $o^*$ whether each legislature will have the opportunity to take a vote to break the agreement. I assume that these probabilities are less than one-half and mutually exclusive.\footnote{While this assumption is made for purposes of exposition and tractability, one can think of $o$ and $o^*$ as being determined by events beyond the executives' or legislatures' control that determine whether or not the legislatures will be willing to consider the lobby's request; for example, they would be affected by the occurrence of an economic crisis that diverts the legislature's attention from less-pressing matters.} This allows me to focus on the key interaction between the domestic actors and abstract from the strategic interaction between the lobbies in the two countries.\footnote{When the legislatures in both countries have an opportunity to break the agreement simultaneously, an incentive for each lobby to free ride on the effort of the other makes the optimization problem of the executives more complex as they are faced with the problem of how best to exploit the free-riding dynamic.}

After this, the uncertainty about the the median legislator's identity is resolved. This is modeled as the realization of a random variable $\ta_b$, the value of which all players observe simultaneously. The ``break stage'' concludes with the legislature making a choice to abide by the agreement or to break the agreement.

@@

\begin{figure}
	\begin{center}
		\input{game-tree.tex}
	\end{center}
	\caption{Extensive Form\label{fig:ext}}
\end{figure}

\subsection{Preferences}
\label{sec:pref}
As the economy is fully separable and the economic and political structures are symmetric, I focus throughout the paper on the home country and the $X$-sector. The details are analogous for $Y$ and foreign.\footnote{These are the relevant sectors for analysis because they are the sectors in which there are lobbies.}

I specify the welfare function of the home legislature as
\begin{equation}
  W_{\mathit{ML}} = \mathit{CS}_X(\tau) + \ga(e,\ve) \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
  \label{eq:ml}
\end{equation}
where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{Labor income $l$ could also be included in both median legislator and executive welfare. I omit it because its inclusion alters none of the results and only serves to complicate the exposition.} I model the decisions of the legislature as being taken by a median legislator (more on this in Section~\ref{sec:choices}). Notice that the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.

Given its expectations and the legislature's preferences, the home lobby chooses its lobbying effort ($e_b$ to influence the break/ratification decision, which for simplicity, I will refer to henceforth simply as ``break,'' and $e_n$ to influence the trade war tariff) to maximize the welfare function:
\begin{equation}
  \expect \left[U_L \right] = \Pr\left[ \text{TradeWar}(\bta,e_b,e_n) \right] \left[ \pi(\tau^{\mathit{n}}) - e_n \right] + \Pr\left[ \text{TradeAgreement}(\bta,e_b,e_n) \right] \pi(\tau^a) - e_b
  \label{eq:lv}
\end{equation}
where $\pi(\cdot)$ is the current-period profit and $\tau^a \ (\tau^\mathit{n})$ is the home country's tariff on the import good under a trade agreement (war). I use the convention throughout of representing a vector of tariffs for both countries $(\tau,\tau^*)$ as a single bold $\bm{\tau}$.

In the first stage of the game, the executives choose the trade agreement tariffs $\bta=\left(\tau^a,\tau^{*a} \right)$ via a negotiating process that I assume to be efficient. In this symmetric environment, it therefore maximizes the expected joint payoffs:
\begin{multline}
  \expect \left[W_E(\bta) \right] + \expect \left[ W_E^{*}(\bta) \right] = \Pr\left[ \text{TradeWar}(\bta) \right] \left[W_E(\bm{\tau}^{\mathit{n}}) + W_E^*(\bm{\tau}^{\mathit{n}}) \right] + \\ \Pr\left[ \text{TradeAgreement}(\bta) \right] \left[W_E(\bta) + W_E^*(\bta) \right]
  \label{eq:jv}
\end{multline}

@@

The home executive's welfare is specified as follows:
\[
  W_E = \mathit{CS}_X(\tau) + \ga_E \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
\]
Note that this is identical to the welfare function for the legislature aside from the weight $\ga_E$ the executive places on the profits of the import industry, which is not a function of lobbying effort. Section~\ref{sec:choices} contains an in-depth discussion of these modeling choices.

\subsection{Information and Equilibrium Selection}
\label{sec:info}
I examine a simple class of equilibria that have the following features.

First, on information. Although uncertainty is present, information about it is symmetric. However, in line with the literature, I assume the lobby's contribution is not observable to the foreign legislature. The implication is that the lobby can directly influence only the home legislature, and so the influence of one country's lobby on the other country's legislature occurs only through the tariffs selected.\footnote{cfr. Grossman and Helpman (1995b), page 685.} Given the game's structure and the assumption that lobbying effort is not observable across international borders, players in the same country can take advantage of more information than those who are in different countries. That is, the solution concept employed here is perfect public equilibrium (PPE).

Second, whenever there is a possibility of multiple equilibria, I will focus on the one that maximizes the welfare of the executives.

Finally, I assume that an external authority can ensure enforcement of the agreement in the one-shot game analyzed here.\footnote{Buzard (2015) considers a similar model in a repeated-game context and can therefore remove this restriction.}

@@

In modeling the objective of the legislature (see Equation~\ref{eq:ml}), I have introduced a variant of the general political support function introduced by Hillman (1982). While Hillman is credited with first proposing that policy makers choose tariffs in order to maximize their own political support, the PFS model contributed the explicit specification that lobbying contributions enter the political support function with constant marginal utility as $C + aW$.

@@

\noindent All proofs are in the Appendix.
